EldersAIR initial PM data work

### 2020-08-21 dustrak_data_cleaning.Rmd
PM data cleaning

---
title: "EldersAIR initial PM data work"
author: "Ethan Walker"
date: "Started 12 Feb 2020, Updated 12 Aug 2020"
output: pdf_document
---

@@

# Histogram for distribution of PM by individual ID
```{r}
# Load data 
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))

@@

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### EldersAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# Load data 
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds"))

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "19", location = "NN", visit = 2)
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)
# filter data for a single home
check_data <- kids_pm %>% 
  filter(area == "NN" & home_winter_id == "10") %>% 
  filter(pm < 0)
```

@@

pm_sample_times_comparison <- read_xlsx(paste0(file_path, "Input/pm_sample_times_comparison.xlsx")) %>% 
  select(area, home, winter_id, sampling_visit, cleaning_comments, 
         next_steps, sensitivity_analysis) %>% 
  mutate(winter_id = as.factor(winter_id),
         sampling_visit = as.factor(sampling_visit))

elders_ids_linked <- read_rds(paste0(file_path, "Output/elders_ids_linked.rds")) %>% 
  filter(winter_id == 1) %>% 
  select(area, home, winter_id, home_winter_id, home_id, treatment) %>% 
  distinct(home, .keep_all = TRUE)

elders_pm_new <- elders_pm %>% 
  mutate(pm = if_else(pm > 10000, 999999, pm)) %>% 
  replace_with_na(replace = list(pm = 999999)) %>% 
  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
  mutate(pm = if_else(pm < 1, 0.5, pm)) %>% 
  #group_by(area, home, home_winter_id, sampling_visit) %>% 
  mutate(time_diff = (datetime_pm - start_datetime_pm)/60,
         pm_datetime_new = if_else(time_diff > 30 & time_diff < 90, 
                                   datetime_pm - 3600, datetime_pm),
         pm_datetime_new = if_else(time_diff < -30 & time_diff > -90, 
                                   pm_datetime_new + 3600, pm_datetime_new)) %>% 
  # make specific time changes
  mutate(pm_datetime_new = if_else(home == "FDE400" & sampling_visit == 1,
                                   pm_datetime_new + 3600, pm_datetime_new),
         pm_datetime_new = if_else(home == "FDE406" & sampling_visit == 1,
                                   pm_datetime_new - 33030*60, pm_datetime_new),
         pm_datetime_new = if_else(home == "NPE5105" & sampling_visit == 1,
                                   datetime_pm + 3600, pm_datetime_new),
         sampling_visit = as.numeric(sampling_visit),
         sampling_visit = if_else(home == "NPE5303" & sampling_visit == 1, 2, sampling_visit),
         sampling_visit = if_else(home == "NPE2404" & sampling_visit == 1, 2, sampling_visit),
         home = if_else(home == "NPE2404", "NPE5204", home),
         pm_datetime_new = if_else(home == "NPE5204" & sampling_visit == 2,
                                   datetime_pm, pm_datetime_new),
         sampling_visit = as.factor(sampling_visit)) %>% 
  arrange(area, home, winter_id, sampling_visit) %>% 
  select(home, area, winter_id, sampling_visit, pm, file_type, day_of_week,
         pm_datetime_new, first_datetime, dtid, pm_total_observations, 
         pm_sample_interval, expected_obs, actual_vs_expected, 
         actual_vs_expected_diff, pm_mean, pm_mean_daily, pm_sampling_day) %>% 
  left_join(pm_sample_times_comparison, c("area", "home", "winter_id", "sampling_visit")) %>% 
  left_join(elders_ids_linked, c("area", "home", "winter_id")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id, sampling_visit)

#write_rds(elders_pm_new, paste0(file_path, "Output/elders_pm_clean.rds"))
```

>>>>

---
title: "EldersAIR initial PM data work"
author: "Ethan Walker"
date: "Started 12 Feb 2020, Updated 21 Aug 2020"
output: pdf_document
---

@@

# Histogram for distribution of PM by individual ID
```{r}
# Load data 
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_new.rds"))

@@

# Run function from above to look at results
## Function run for each household to check for odd trends/gaps in the data
## Notes from this process on uploaded to Box:
### EldersAIR > Ethan - data cleaning and analysis > pm_cleaning_summary
```{r}
# Load data 
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")
elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_new.rds"))

# Change function input to look at specific home
# Run these 4 lines together
pm_time_plot_function(location = "NPT", id = "52", visit = 1)
head(gaps, 10)
head(pm_desc, 10)
tail(pm_desc, 10)

# filter data for a single home
check_data <- elders_pm %>% 
  filter(area == "NN" & home_winter_id == "11" & sampling_visit == 1) %>% 
  filter(pm < 0)
```

@@

# Load file with data cleaning notes
pm_sample_times_comparison <- read_xlsx(paste0(file_path, "Input/pm_sample_times_comparison_21aug2020.xlsx")) %>% 
  select(area, home, home_winter_id, winter_id, sampling_visit, cleaning_comments, 
         next_steps, sensitivity_analysis) %>% 
  mutate(winter_id = as.factor(winter_id),
         home_winter_id = as.factor(home_winter_id),
         sampling_visit = as.factor(sampling_visit))

elders_pm_new <- elders_pm %>% 
  mutate(pm = if_else(pm > 12000, 999999, pm)) %>% 
  replace_with_na(replace = list(pm = 999999)) %>% 
  # join with file on data cleaning notes - refer to pm_sample_times_comparison file
  left_join(pm_sample_times_comparison, c("area", "home", "home_winter_id", "winter_id", "sampling_visit")) %>% 
  # filter out sampling visits with bad data - refer to pm_sample_times_comparison file
  filter(next_steps != "Do not use due to short run time") %>% 
  filter(next_steps != "Do not use due to high baseline/minimum values") %>% 
  filter(next_steps != "Do not use due to negatives") %>% 
  # replace PM < 1 (lower DustTrak limit) with 0.5 (lower limit / 2) 
  # only doing this for files with a few negative values
  mutate(pm = if_else(pm < 1, 0.5, pm)) %>% 
  # fix files with incorrect datetimes 
  mutate(datetime_pm = ymd_hms(datetime_pm),
         pm_datetime_new = if_else(next_steps == "Subtract 1 hour from datetime", 
                                   datetime_pm - 3600, datetime_pm),
         pm_datetime_new = if_else(next_steps == "Add 1 hour to datetime", 
                                   datetime_pm + 3600, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Add 1 day to datetime", 
                                   datetime_pm + 86400, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Add 2 hours to datetime", 
                                   datetime_pm + 7200, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Add 174970 hours to datetime", 
                                   datetime_pm + 629892000, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Subtract 2 hours from datetime", 
                                   datetime_pm - 7200, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Subtract 4 hours from datetime", 
                                   datetime_pm - 14400, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Subtract 676 hours from datetime", 
                                   datetime_pm - 2433600, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Subtract 7.75 hours from datetime", 
                                   datetime_pm - 27900, pm_datetime_new),
         pm_datetime_new = if_else(next_steps == "Subtract 78 hours from datetime", 
                                   datetime_pm - 280800, pm_datetime_new)) %>% 
  # filter(next_steps == "Subtract 78 hours from datetime")
  arrange(area, home, winter_id, sampling_visit, pm_datetime_new) %>% 
  ungroup()

# write_rds(elders_pm_new, paste0(file_path, "Output/elders_pm_clean.rds"))
```
